Sequential rank agreement methods for comparison of ranked lists

### 2015-06-03 latex/paper.Rnw
sqrt result and plot label

Sequential rank agreement can be generalized to incomplete/censored
lists in the following way.  Let $\Lambda_l, l=1, \ldots, L$ be the
set of items found in list $l$ so $\Lambda_l$ is the top $k_l$ list of
items from list $l$ where $k_l = |\Lambda_l|$. Note that if
$k_1=\cdots=k_L=k$ then we observe the top $k$ items for each of the
$L$ lists. For censored lists the rank function becomes
\begin{equation}
\tilde R_l(X_p) = \left\{\begin{array}{cl} \{\pi_l^{-1}(p)\} & \text{ for } p\in \Lambda_l \\
\{k_l+1,\dots,P\} & \text{ for } p \not\in \Lambda_l\end{array}\right.
\end{equation}
where we only know that the rank for the unobserved items in list $l$
must be larger than the largest rank observed in that list.
Unless we have additional prior knowledge about the distribution of the
unranked items then we can restrict attention of the sequential rank
agreement to
\begin{equation}
d \leq \max(k_1, \ldots, k_L),
\end{equation}
since we have no information about the ordering of the items after a
depth $d$ where all lists are censored.

The agreement, $A(X_p)$, cannot be computed directly for all
predictors because we only observe the censored version of $R(X_p)$
for some of the lists. Instead we assume that the rank assigned to
item $X_p$ in list $l$ is uniformly distributed among the ranks that
have \emph{not} been assigned for list $l$. The algorithm is described
in detail in Algorithm~\ref{sra-algorithm}.

%, so the augmented agreement
%becomes
%\begin{equation}
%\bar{A}(X_p) = \frac{\sum_{r_1; r_1\in \tilde R_1(X_p)}  \cdots
%  \sum_{r_L; r_L\in \tilde R_L(X_p)} A(X_p)}{\prod_l |\tilde
%  R_l(X_p)|}. \label{formula:censored-agreement}
%\end{equation}
%Similarly, the sequential rank agreement becomes
%\begin{equation}
%XXX \textrm{SRA}(d)= \sqrt{\frac{\sum_{\{p \in S_d\}}(L-1)\tilde{A}(X_p)^2}{(L-1)|S_d|}}, \label{def:sra-censored}
%\end{equation}

The rankings within a single list are clearly not independent since
each rank must appear exactly in each list. Thus, by randomizing
unobserved items as described in \eqref{formula:censored-agreement}

@@

@@

<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE >>=
ysize <- 80
plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
lines(res2[1:ysize], lwd=2, col="purple")
##
# col1 <- makeTransparent()
null <- sqrt(random_list_sra(inputmatrix, B=1, n=400))
null3 <- sqrt(random_list_sra(inputmatrix, B=20, n=200))
## matlines(null[1:ysize,], col="gray", lty=1)

@@

The average overlap at depth $d$ is defined as the average of the
overlaps until $d$,
\begin{equation}
AO(d) = \sum_{i=1}^d \frac{O(d)}{d}.
\end{equation}
This definition indirectly ensures that extra emphasis is included on
the top of the list since the items

@@

<<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE>>=
mmyres <- sqrt(sra(inputmatrix[,1:2]))
mmyres2 <- sqrt(sra(inputmatrix[-c(1,2),1:2]))
xvals <- 1:500
plot(mmyres, xlim=c(1, 300), ylim=c(0,1), type="l")
lines(xvals[-c(1,2)], mmyres2[1:498], col="red")
@

>>>>

Sequential rank agreement can be generalized to incomplete/censored
lists in the following way.  Let $\Lambda_l, l=1, \ldots, L$ be the
set of items found in list $l$ so $\Lambda_l$ is the top $k_l$ list of
items from list $l$ where $k_l = |\Lambda_l|$. Note that if
$k_1=\cdots=k_L=k$ then we observe the top $k$ items for each of the
$L$ lists. For censored lists the rank function becomes
\begin{equation}
\tilde R_l(X_p) = \left\{\begin{array}{cl} \{\pi_l^{-1}(p)\} & \text{ for } p\in \Lambda_l \\
\{k_l+1,\dots,P\} & \text{ for } p \not\in \Lambda_l\end{array}\right.
\end{equation}
where we only know that the rank for the unobserved items in list $l$
must be larger than the largest rank observed in that list.

The agreement, $A(X_p)$, cannot be computed directly for all
predictors when we have censored lists because we only do not know the
exact rank assigned to each item in all lists.  Instead we assume that
the rank assigned to an item $X_p$ that is not present in list $l$
(i.e., is censored) is uniformly distributed among the ranks that have
\emph{not} been assigned for list $l$.  The rankings within a single
list are clearly not independent since each rank must appear exactly
in each list. Thus we cannot simply assign

Thus, by randomizing unobserved items as described in
\eqref{formula:censored-agreement}

The algorithm is described in
detail in Algorithm~\ref{sra-algorithm}.

@@

Unless we have additional prior knowledge about the distribution of the
unranked items then we can restrict attention of the sequential rank
agreement to
\begin{equation}
d \leq \max(k_1, \ldots, k_L),
\end{equation}
since we have no information about the ordering of the items after a
depth $d$ where all lists are censored.

@@

<<fig2b, fig=TRUE,echo=FALSE,include=FALSE,cache=TRUE >>=
ysize <- 80
plot(res[1:ysize], lwd=3, col="red", type="l", ylim=c(0,1300), ylab="Sequential rank agreement", xlab="Depth")
lines(res2[1:ysize], lwd=2, col="purple")
##
# col1 <- makeTransparent()
null <- sqrt(random_list_sra(inputmatrix, B=1, n=40))
null3 <- sqrt(random_list_sra(inputmatrix, B=20, n=20))
## matlines(null[1:ysize,], col="gray", lty=1)

@@

The average overlap at depth $d$ is defined as the average of the
overlaps until $d$,
\begin{equation}
AO(d) = \sum_{i=1}^d \frac{O(d)}{d}.
\end{equation}
This definition directly ensures that extra emphasis is included on
the top of the list since the overlap of the items towards the top are
included in the majority of the calculations.

@@

<<fig-ao2, echo=FALSE, fig=TRUE,include=FALSE>>=
mmyres <- sqrt(sra(inputmatrix[,1:2]))
mmyres2 <- sqrt(sra(inputmatrix[-c(1,2),1:2]))
xvals <- 1:500
plot(mmyres, xlim=c(1, 300), type="l")
lines(xvals[-c(1,2)], mmyres2[1:498], col="red")
@
